---
title: "Visualizing Text and Distributions"
author: "Katie Dills"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# Data Visualization Project 03


# Text Mining

## Top 100 Songs

I decided to use the text mining techniques I have learned over the past few semesters to examine the lyrics of a few of the top hit songs.


```{r}
library(tidyverse)
library(tidytext)

billboard_top100 <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/BB_top100_2015.csv")

head(billboard_top100)
```

***

#Word Count

First I wanted to take a look at the word count to better understand what words were used the most and who the artist of the songs were. The great thing about reviewing top hit songs with the highest word count for one particular song is that even though the name of the song is not listed I immediatley knew the first two songs.

```{r}
lyrics_tokens <- billboard_top100 %>% 
  unnest_tokens (word, Lyrics) %>%
  anti_join(stop_words, by = "word") %>%
  group_by(Artist) %>% 
  count(word, sort = TRUE) %>% 
  top_n(9, n) %>% 
  ungroup() %>% 
  mutate(word = fct_inorder(word))

head(lyrics_tokens)
```

I then decided to chart any words that were used by an artist of **25** times in a single song to see if any words were popular among multiple artist. As you can see in the graph below love is a very popular word for the billboard songs.

```{r}
lyrics_tokens %>% 
  filter(n >= 25) %>% 
ggplot(aes(x = n, y = fct_rev(word), fill = Artist)) +
  geom_col() +
  guides(fill = FALSE) +
  labs(x = NULL, y = NULL) +
  scale_fill_viridis_d()+
  labs(title = "Highest Word Count",
       subtitle = "(Highest word count per song)",
       x = "Word Count", 
       y = "Word") +
  theme_minimal()
```

***

#Term Bigrams

I then decided to explore most used bigrams of the lyrics. I didn't care for the top word phrases as most were the same word being repeated which is common in songs. But I deicided against graphing these.

```{r}
lyrics_bigrams <- billboard_top100 %>% 
  unnest_tokens(bigram, Lyrics, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>% 
  filter(!word1 %in% stop_words$word) %>% # remove stopwords
  filter(!word2 %in% stop_words$word) %>% # remove stopwords
  unite(bigram, word1, word2, sep = " ") %>% 
  group_by(Artist) %>% 
  count(bigram, sort = TRUE) %>% 
  top_n(9, n) %>% 
  ungroup() %>% 
  mutate(bigram = fct_inorder(bigram))

head(lyrics_bigrams)

```
***

#Term Sentiments

I looked at the sentiments of the lyrics, specifically whether the words are negative or positive. There was a wide range of net sentiment's however there seemed to be a lot more negative sentiments than positive. I picked a few sample artist and graphed their net sentiment to see how they compared with positive and negative sentiments from their songs. Once they were graphed, there appears to be more positive net sentiment songs than negative. However interestingly **Taylor Swift** had a song that was very negative, but it's not surprising considering she does come out with a large number of break up songs.

```{r}
lyrics_sentiment <- billboard_top100 %>% 
  unnest_tokens(word, Lyrics) %>%
  group_by(Artist) %>% 
  mutate(word_count = 1:n(),
         index = word_count %/% 500 + 1) %>% 
  inner_join(get_sentiments("bing")) %>%
  count(Artist, index = index, sentiment) %>% 
  pivot_wider(names_from = sentiment, values_from = n) %>%
  mutate(net_sentiment = positive - negative)

print(lyrics_sentiment)
```

```{r}
top_artist_sentiment <- filter(lyrics_sentiment, Artist %in% c("the weeknd", "meghan trainor", "rihanna", "maroon 5", "beyonce", "drake", "nick jonas", "taylor swift", "sam smith"))


head(top_artist_sentiment)
```


```{r}
ggplot(top_artist_sentiment, 
       aes(x = index, y = net_sentiment, fill = net_sentiment > 0)) +
  geom_col() +
  guides(fill = FALSE) +  
  labs(x = NULL, y = "Net sentiment") +
  scale_fill_manual(name = "", labels = c("Positive", "Negative"),
                     values = c("#FF851B", "#3D9970")) +
  facet_wrap(vars(Artist), scales = "free_x") +
  labs(title = "Song Sentiment",
       subtitle = "(Sample Artist Song Sentiment)",
       x = "Index", 
       y = "Net Sentiment") +
  theme_minimal()
```

***

#Term Frequency

Lastly I reviewed the lyrics frequency however I didn't like how the term frequencies appeared so I decided not to graph this set either.

```{r}
lyrics_words <- billboard_top100 %>% 
  unnest_tokens(word, Lyrics) %>% 
  count(Artist, word, sort = TRUE) %>% 
  ungroup()

lyrics_tf_idf <- lyrics_words %>% 
  bind_tf_idf(word, Artist, n) %>%
  arrange(desc(tf_idf))

lyrics_tf_idf_plot <- lyrics_tf_idf %>% 
  group_by(Artist) %>% 
  top_n(8) %>% 
  ungroup() %>% 
  mutate(word = fct_inorder(word))

head(lyrics_tf_idf_plot)
```

***

# Final Thoughts

The top billboard song lyrics made great data sets for text mining purposes. The most interesting aspects I found in this dataset were the word with the highest count from songs and the net sentiments of the songs. The word count was interesting as the songs are very obvious from of key word alone, such as the song *Shake It Off*. Also the song sentiments are interesting as there are both negative and positive sentiments across the top billboard songs. Also the topics that the artist tend to sing about show in the sentiments as well, for example Taylor Swift having a song with a large amount of negative sentiments. 


***
